DOpElib: Differential Equations\\ and Optimization Environment

### 2013-06-07 Examples/PDE/StatPDE/Example3/content.tex
Preparation for moving the examples to a new ordering.

\textbf{General problem description}

\vspace{0.2cm}

This example is an extension the previous one.
We solve an stationary FSI problem either with 
INH material (see Problem definition before) 
or St. Venant Kirchhoff material STVK: \index{structure} \index{structure!compressible St.Venant-Kirchhoff (STVK) material}
\begin{Problem}[Compressible Saint Venant-Kirchhoff, Lagrangian framework]
  \label{prob:stvk:lagrange}
  Find $\{\hat v_f,\hat u_f\} \in \{ \hat v_f^D + \hat V\} 
  \times \{ \hat u_f^D + \hat V\}$, 
  such that 
  \begin{eqnarray}
    \begin{aligned}   
      (\hat J_s \hat\sigma_s \hat F_s^{-T},\hat\nabla\hat\phi^v)_{\hat\Omega_s}
      &= \langle\hat J_s\hat n_s\cdot \hat g_s^\sigma\hat
      F_s^{-T},\hat\phi^v  \rangle_{\hat\Gamma_i}
      &&\quad\forall\hat \phi^v\in\hat V_s\\
      %%%%%%%%%%%%%%%% 
      (\hat v_s,\hat\phi^u)_{\hat\Omega_s}&=0
      &&\quad\forall\hat \phi^u\in\hat V_s    ,\\
    \end{aligned}
  \end{eqnarray}
  where $\rho_s$ is the density of the structure, $\mu_s$ and $\lambda_s$ 
  the Lam\'e coefficients, $\hat n_s$ the outer normal vector at $\hat\Gamma_i$, 
  $\hat g_s^\sigma$ some forces on the interface.\index{Lam\'e coefficients} The properties of the STVK material
  is specified by the constitutive law
  \begin{eqnarray}
    \hat\sigma_s:=\hat J^{-1} \hat F (\lambda_s (tr\hat E) I + 2\mu_s \hat E)\hat F^{-T}.
  \end{eqnarray}
\end{Problem}
Often, the elasticity properties of structure materials is characterized by 
Poisson's ratio $\nu_s$ ($\nu_s < \frac{1}{2}$ \index{Poisson's ratio} for 
compressible materials) and the Young modulus $E$ \index{Young modulus}. The 
relationship to the Lam\'e coefficients $\mu_s$ and $\lambda_s$ is
given by:
\begin{eqnarray}
\nu_s = \frac{\lambda_s}{2(\lambda_s + \mu_s)}, \qquad 
E = \frac{\mu_s (\lambda_s + 2\mu_s)}{(\lambda_s + \mu_s )} . 
\end{eqnarray}
The whole equation system is solved on the benchmark configuration domain. For 
details on parameters and geometry, we refer to the 
numerical FSI benchmark \index{fluid-structure interaction (FSI)} \index{fluid-structure interaction (FSI)!FSI benchmark} proposal from Hron and Turek [2006].

The code is established by computing the stationary FSI benchmark example FSI 1 with 
the following values of interest: $x$-displacement, $y$-displacement, drag, and lift.\index{functional} \index{functional!drag} \index{functional!lift} \index{functional!deflection}

\vspace{0.2cm}

\textbf{Program description}

\vspace{0.2cm}

Compared to the two former examples, there are some differences which we will briefly discuss in the following. First of all, the problem is nonlinear in contrast to the former ones. We work on a different domain (given in the \textit{benchfst0100tw.inp} file), namely a channel with a cylinder put at half height near the inflow boundary; further \textit{.inp} files yield the possibility to vary the domain.\\
Furthermore, in the \textit{dope.prm} parameter file there are two additional subsections which are added only for the current problem. From the denotation of these subsections one can immediately see where in the code the parameters are used.\\
As we want to compute certain benchmark quantities, we have to regard corresponding functionals in the \textit{functionals.h} file. The pressure at a point as well as the displacement in $x$- and $y$-directions are point values; furthermore we implement the drag and lift functionals (for which we need the additionally defined problem parameters).\\
As before, we build up the cell and boundary equations and matrices in the \textit{localpde.h} file. Apart from using the additionally defined problem parameters and modelling compressible STVK material instead of INH material (which leads to changes in the weak formulation of the equations), there are no major differences to the corresponding file in the last example.\\
In the \textit{main.cc} file, we have to include additional header files from the deal.II library concerning error estimation and grid refinement. Further on, everything is pretty much the same as in the last example, but we have to use the \texttt{SetBoundaryFunctionalColors} function of the \texttt{PDEProblemContainer} class to be able to compute drag and lift in the respective functional classes in \textit{functionals.h}.\\
The main innovation in contrast to the preceding examples is the refinement of the grid combined with a simple error estimator given in the deal.II \texttt{KellyErrorEstimator} class. If we look at the output of our program, everything is computed several times (once on each refinement level). Furthermore, we see that several Newton steps are needed on each refinement level; this is due to the nonlinearity of the current problem.

>>>>

\subsubsection{General problem description}
\index{stationary PDE}
In this example we consider the same setting as in subsection~\ref{PDE_Stat_Stokes}, the only
difference is that we want to employ the hp-Finite-Elements. So the equation we solve is still 
the stationary incompressible Stokes equation \index{stationary PDE!Stokes equation}. Here,
we use the symmetric stress tensor which has a little consequence when using 
the do-nothing outflow condition. In strong formulation we have
\begin{align} 
-\frac{1}{2}\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p &= f \\ 
\nabla \cdot v &= 0 \notag
\end{align} 
on the domain $\Omega = [-6,6]\times [0,2]$.  We split $\partial \Omega = \Gamma_D \cup \Gamma_{out}$. The right hand side of the channel is $\Gamma_{out}$ on which we describe the free outflow condition, on the rest of the boundary we prescribe dirichlet values (An parabolic inflow on the left hand side and zero on the upper and lower channel walls). We choose for simplicity $f=0$..\\

\subsubsection{Adding hp-Elements}
One sees by comparing  the \texttt{main.cc}-file of this problem with the one of subsection~\ref{PDE_Stat_Stokes} that the change to hp-Elements is really easy, so we will keep the description short. In comparison to example \ref{PDE_Stat_Stokes} the \textit{localpde.h} and \textit{functionals.h} have not changed, but we have one additional file, namely \textit{indexsetter.h}, in which the class \texttt{ActiveFEIndexSetter} is defined.

In the hp-framework we have a stack of finite elements (a \texttt{hp::FECollection}) given. We assign each cell of the triangulation an fe-index which determines which finite element we use on this cell. The \texttt{ActiveFEIndexSetter} class manages these indices, see there for more information.

The changes in \textit{main.cc} are also minimal and are highlighted in the source code. Obviously, we use \texttt{FECollection} and \texttt{QCollection} as well as a different DoFHandler.
\begin{verbatim}
  	#define DOFHANDLER hp::DoFHandler
  	#define FE hp::FECollection
  	...
  	typedef hp::QCollection<DIM> QUADRATURE;
  typedef hp::QCollection<DIM - 1> FACEQUADRATURE;
\end{verbatim}
Apart from that we have only to tell the space time handler the distribution of the finite element indices:
\begin{verbatim}
  ActiveFEIndexSetter<2> indexsetter(pr);
  STH DOFH(triangulation, state_fe_collection, indexsetter);
\end{verbatim}

### 2013-06-07 Examples/PDE/StatPDE/Example8/content.tex
Preparation for moving the examples to a new ordering.

Laplace in 3D
on locally refined grids with Kelly estimator
with iterative solver (CG without preconditioning)

>>>>

\subsubsection{General problem description}

Similar to the previous example, we consider the following benchmark problem from plasticity theory:
\begin{equation}
   (\Pi(\sigma(u)),\varepsilon(\varphi)) = (g,\varphi)_{\Gamma_N}.
\end{equation}
Here $\tilde{\Omega}$ is again the quadratic domain with a circular hole around the center cut out. Again, we restrict our actual computational domain $\Omega$ to the upper left quarter of $\tilde{\Omega}$ for reasons of symmetry.\\
We use the symmetric strain tensor $\varepsilon(v) := \frac{1}{2}(\nabla v + \nabla v^T)$, and the symmetric stress tensor $\sigma$ is defined as
\begin{equation*}
   \sigma(v) := 2\mu \varepsilon(v)^D + \kappa \cdot tr (\varepsilon(v)) I
\end{equation*}
where $\tau^D$ is the deviatoric part of a tensor $\tau$, in two dimensions defined as
\begin{equation*}
   \tau^D := \tau - \frac{1}{2} tr(\tau) I.
\end{equation*}
The main difference with respect to the elastic case is the projection operator $\Pi$ in equation $(1)$. It is defined as follows:
\begin{equation*}
\Pi(\tau) = \left\{
            \begin{array}{lr}
            \tau & |\tau^D| \leq \sigma_0\\
            \sigma_0 |\tau^D|^{-1} \tau^D + \frac{1}{2} tr(\tau) I & |\tau^D| > \sigma_0\\ 
            \end{array}
            \right. 
\end{equation*}
In our computations, we choose $\sigma_0 = \sqrt{\frac{2}{3}}\cdot 450$, and the above parameters $\mu$ and $\kappa$ as $\mu = 80193.800283$ resp. $\kappa = 190937.589172$. The corner points of our computational domain are the same as before, and the boundary conditions are not altered, either.\\
The goal of our computations is to detect a subdomain in $\Omega$ where plastic behaviour occurs (compare \textit{E. Stein (editor), Error-controlled Adaptive Finite Elements in Solid Mechanics, Wiley (2003), pp. 386 - 389}). This subdomain depends on the righthand side $g$ in equation $(1)$ which we write as $g = \lambda \cdot p$ with $p = 100$ and $\lambda \in [1.5;4.5]$.\\

\subsubsection{Program description}

The code of the current example is nearly identical to the code of the previous one. The only difference worth mentioning is the change of the equations which leads to different implementations of the \texttt{CellEquation, CellMatrix} and \texttt{BoundaryEquations} functions in \textit{localpde.h}.\\
Furthermore, the elasticity equations solved in the last example are linear, whereas the plasticity equations are nonlinear; this difference is evident also from the output (here, we need several Newton steps until convergence).\\
The functionals that appear in the output yield additional information and are not required in the above problem setting. The subdomain with plastic behaviour we want to detect can be visualized from the \textit{.vtk} files written to the \texttt{Results/Mesh} subfolders.
